---
title: "DSCI 401 - Homework 1"
author: "Isabel Heard"
date: "09/08/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1.(20points) 
Set up a github account and create a repository called DSCI401

## 2. (40 points) 
Create an R Markdown file and run the following code:
```{r}
#getwd()

df <- read.csv("/Users/isabelheard/Downloads/DSCI 401/DSCI401_git/data/cubs_all_time.csv")
head(df)
```

## Answer the following questions using R:
# (a) 
How many total games have the Cubs won and lost between 1876 and 2022?
```{r}
sum(df$W)
sum(df$L)
```
11141 games won between 1876 and 2022.
10592 games lost between 1876 and 2022.


# (b) 
What year did the Cubs score the most runs? 
```{r}
which.max(df$R)
df[129, ]

max(df$R)
```
The cubs scored the most runs (1056) in 1894.


What year did the Cubs score the fewest runs? 
```{r}
which.min(df$R)
df[3, ]

min(df$R)
```
The cubs scored the least runs (265) in 2020.


Do you have any thoughts about the year that the Cubs scored the fewest runs?
```{r}

```
Covid-19.

# (c) 
In how many seasons was the Cubs total attendance (i.e. the variable Attendance) over 3 million?
```{r}
df$Attendance >= 3000000

length(which(df$Attendance >= 3000000))
```
2 seasons the total club attendance was over 3 million.
