---
title: "DSCI 401 - Homework 1"
author: "Isabel Heard"
date: "09/08/2023"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1.Set up a github account and create a repository called DSCI401
https://github.com/iisabel/DSCI-401.git

# 2. Create an R Markdown file and run the following code:
```{r}
df <- read.csv("/Users/isabelheard/Downloads/DSCI 401/DSCI401_git/data/cubs_all_time.csv")
#head(df)
```

## A) How many total games have the Cubs won and lost between 1876 and 2022?
```{r}
sum(df$W)
sum(df$L)
```
11141 games won between 1876 and 2022.

10592 games lost between 1876 and 2022.


## B) What year did the Cubs score the most runs? 
```{r}
#max(df$R)
which.max(df$R)
df[129, ]
```
The cubs scored the most runs (1056) in 1894.


## B2) What year did the Cubs score the fewest runs? 
```{r}
#min(df$R)
which.min(df$R)
df[3, ]
```
The cubs scored the least runs (265) in 2020.


## B3) Do you have any thoughts about the year that the Cubs scored the fewest runs?

COVID-19 meant that there were fewer games.

## C) In how many seasons was the Cubs total attendance (i.e. the variable Attendance) over 3 million?
```{r}
#df$Attendance >= 3000000
length(which(df$Attendance >= 3000000))
```
12 seasons the total club attendance was over 3 million.


# Python code link
https://colab.research.google.com/drive/1Pg0RwYnDq1VCvyB0_iPm5WyZk1lQ98Cy?usp=sharing

