---
title: "DSCI 401 HW 4"
author: "Isabel Heard"
date: "10/20/2023"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1

## A
Use the HELPrct data from the mosaicData to calculate the mean of all numeric variables (be sure to exclude missing values).
```{r}
library(mosaicData)
library(tidyverse)
library(dplyr)
data(HELPrct)
#head(HELPrct)
#summary(HELPrct)

HELPrct %>%
  summarize(across(where(is.numeric), mean, na.rm = TRUE))
```


## B
Find the mean of all the numeric variables stratified by sex and age group where age groups are defined as ranges of 10 years (i.e. 0-10, 10-20, 20-30, etc).
```{r}
#Age groups and labels
age_groups <- seq(0, max(HELPrct$age), by = 10)
age_labels <- paste(age_groups, age_groups + 10, sep = "-")
#Add a label for ages greater than the maximum age - helps with error
age_labels[length(age_labels)] <- paste(age_labels[length(age_labels)], "and above")
df <- HELPrct %>% mutate(age_group = cut(age, breaks = c(age_groups, Inf), labels = age_labels, include.lowest = TRUE))

#Mean for each numeric variable, stratified by sex and age group
result <- df %>%
  group_by(sex, age_group) %>%
  summarize(across(where(is.numeric), mean, na.rm = TRUE)) #same code from previous problem
print(result)
```


## C
Using the data set created in the previous problem, create a set of line plots with the average age of the age group on the x-axis and each of other numeric variables on the y-axis in separate plots stratified by sex. (Note: You are not allowed to use a for loop here or simply copy-and- paste 20 times!)
```{r}
library(ggplot2)
df_long <- result %>%
  pivot_longer(cols = c("anysubstatus", "cesd", "d1", "daysanysub", "dayslink", "drugrisk", "e2b", "female", "i1", "i2", "id", 
                        "indtot", "linkstatus", "mcs", "pcs", "pss_fr", "sexrisk", "avg_drinks", "max_drinks", "hospitalizations"), names_to = "variable")

plots <- df_long %>%
  ggplot(aes(x = age_group, y = value, group = interaction(variable, sex), color = sex, linetype = sex)) +
  geom_line() +
  facet_wrap(vars(variable), scales = "free_y", ncol = 4, nrow = 5) +
  labs(x = "Age Group", y = "Numeric Variable", title = "Average Age vs. Given Numeric Variable") +
  scale_linetype_manual(values = c("solid", "solid")) +  
  scale_color_manual(values = c("blue", "red")) +        
  theme_minimal()
print(plots)

```


# 2
The team IDs corresponding to Brooklyn baseball teams from the Teams data frame from the Lahman package are listed below. Use map int() to find the number of seasons in which each of those teams played by calling a function called count seasons.
```{r}
library(Lahman)
data(Teams)

#List of Brooklyn baseball teams
bk_teams <- c("BR1", "BR2", "BR3", "BR4", "BRO", "BRP", "BRF")

#Count seasons by teamID/yearID
count_seasons <- function(team_id) {
  Teams %>%
    filter(teamID == team_id) %>%
    distinct(yearID) %>%
    nrow()}

#Map to count seasons for each team
season_counts <- map_int(bk_teams, count_seasons)

#Make df
result_df <- data.frame(teamID = bk_teams, seasons_played = season_counts)
print(result_df)
```


# Colab Link
https://colab.research.google.com/drive/1Yx9utHdcMaggM7X69Yn5h0Jm9ljwOHVC?usp=sharing

